### 2020-07-19 paper.tex
making main algorithm box less cluttered

\begin{algorithm*}[h]

@@

    \State $(y, \log q_x(y)) \gets \textproc{trace-and-score}(\mathcal{Q}_x)$ \Comment{Sample $y \sim q_x(\cdot)$; compute log-density via probabilistic program}
    \State $(x' \oplus y', \log D) \gets \textproc{run-involution}(\mathcal{F}, x \oplus y)$ \Comment{{\small Compute $(x' \oplus y') = f(x \oplus y)$ and log(Radon-Nikodym derivative)}}
    \State $\log \tilde{p}(x \oplus b) \gets \textproc{score}(\mathcal{P}, x \oplus b)$ \Comment{Compute log-density via probabilistic program}
    \State $\log \tilde{p}(x' \oplus b) \gets \textproc{score}(\mathcal{P}, x' \oplus b)$ \Comment{Compute log-density via probabilistic program}
    \State $\log q_{x'}(y') \gets \textproc{score}(\mathcal{Q}_{x'}, y')$ \Comment{Compute log-density via probabilistic program}
    \State $\alpha \gets \min\{1, \exp(\log \tilde{p}(x' \oplus b) - \log \tilde{p}(x \oplus b) + \log q_{x'}(y') - \log q_x(y) + \log D)\}$ \Comment{Acceptance probability}

@@

We now introduce a simple differentiable programming language for specifying involutions $f$.
The language needs to have syntax for reading the value from an address in $x \oplus y \in Z$ and writing to an address in $x' \oplus y' \in Z$.
To read a value $(x \oplus y)[a]$ at address $a$, we use the \texttt{@read} keyword:
\noindent
\begin{center}
\begin{tabular}{c}
{\begin{lstlisting}[basicstyle=\small\ttfamily]
    value = @read(<address>, <type>)
\end{lstlisting}}
\end{tabular}
\end{center}
The first argument is the address $a$ and the second argument is either $\mathtt{:}\mathtt{discrete}$ or $\mathtt{:}\mathtt{continuous}$, and informs the interpreter whether the random choice at that address is drawn from a discrete or continuous distribution (this information will be used to support automatic differentiation).
Recall that $x$ is the trace of the model probabilistic program, and $y$ is the trace of the auxiliary probabilistic program.
Each address $a$ therefore needs to specify which of these traces to read from, and the address within that trace.
The traces are given names in the function signature:
\noindent
\begin{center}
\begin{tabular}{c}
{\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
@transform f (model_in, aux_in) to (model_out, aux_out)
begin
..
end
\end{lstlisting}}
\end{tabular}
\end{center}
Here, the traces $x$, $y$, $x'$ and $y'$ are given names $\mathtt{model\_in}$, $\mathtt{aux\_in}$, $\mathtt{model\_out}$, and $\mathtt{aux\_out}$, respectively.
The syntax for address $\mathtt{a}$ within trace $\mathtt{trace}$ is $\mathtt{trace[a]}$.
For example, to read the value of a continuous address $\mathtt{:}\mathtt{r}$ from the input model trace ($x$):
\noindent
\begin{center}
\begin{tabular}{c}
{\begin{lstlisting}[basicstyle=\small\ttfamily]
    val = @read(model_in[:r], :continuous)
\end{lstlisting}}
\end{tabular}
\end{center}
The syntax for writing to an address in $x' \oplus y' \in Z$ is similar.
For example to write a value $\mathtt{val}$ to $x'[\mathtt{:}\mathtt{r}]$:
\noindent
\begin{center}
\begin{tabular}{c}
{\begin{lstlisting}[basicstyle=\small\ttfamily]
    @write(model_in[:r], val, :continuous)
\end{lstlisting}}
\end{tabular}
\end{center}
Note that the input traces $x, y$ are distinct from the output traces $x', y'$;
input traces can only be read from, and output traces can only be written to.
For example, it is not possible to write an output trace and then read the written value from it later.

>>>>

\begin{algorithm*}[h!]

@@

    \State\vspace{-2mm}
    \LineComment{Sample $y \sim q_x(\cdot)$ and compute log-density via probabilistic program $\mathcal{Q}$}
    \State $(y, \log q_x(y)) \gets \textproc{trace-and-score}(\mathcal{Q}_x)$
    \State\vspace{-2mm}
    \LineComment{{\small Compute $(x' \oplus y') = f(x \oplus y)$ and log(Radon-Nikodym derivative) via differentiable program $\mathcal{F}$}}
    \State $(x' \oplus y', \log D) \gets \textproc{run-involution}(\mathcal{F}, x \oplus y)$
    \State\vspace{-2mm}
    \LineComment{Compute log-density via probabilistic program $\mathcal{P}$}
    \State $\log \tilde{p}(x \oplus b) \gets \textproc{score}(\mathcal{P}, x \oplus b)$
    \State\vspace{-2mm}
    \LineComment{Compute log-density via probabilistic program $\mathcal{P}$}
    \State $\log \tilde{p}(x' \oplus b) \gets \textproc{score}(\mathcal{P}, x' \oplus b)$ 
    \State\vspace{-2mm}
    \LineComment{Compute log-density via probabilistic program $\mathcal{Q}$}
    \State $\log q_{x'}(y') \gets \textproc{score}(\mathcal{Q}_{x'}, y')$ 
    \State\vspace{-2mm}
    \LineComment{Compute acceptance probability and accept or reject}
    \State $\alpha \gets \min\{1, \exp(\log \tilde{p}(x' \oplus b) - \log \tilde{p}(x \oplus b) + \log q_{x'}(y') - \log q_x(y) + \log D)\}$ 

@@

We now introduce a simple differentiable programming language for specifying involutions $f$.
The language needs to have syntax for reading the value from an address in $x \oplus y \in Z$ and writing to an address in $x' \oplus y' \in Z$.
To read a value $(x \oplus y)[a]$ at address $a$, we use the \texttt{@read} keyword:
\noindent
\begin{center}
\begin{tabular}{c}
{\begin{lstlisting}[basicstyle=\small\ttfamily]
    value = @read(<address>, <type>)
\end{lstlisting}}
\end{tabular}
\end{center}
The first argument is the address $a$ and the second argument is either $\mathtt{:}\mathtt{discrete}$ or $\mathtt{:}\mathtt{continuous}$, and informs the interpreter whether the random choice at that address is drawn from a discrete or continuous distribution (this information will be used to support automatic differentiation).
Recall that $x$ is the trace of the model probabilistic program, and $y$ is the trace of the auxiliary probabilistic program.
Each address $a$ therefore needs to specify which of these traces to read from, and the address within that trace.
The traces are given names in the function signature:
\noindent
\begin{center}
\begin{tabular}{c}
\begin{lrbox}{\mybox}%
\begin{lstlisting}[basicstyle=\small\ttfamily]
@transform f (model_in, aux_in) to (model_out, aux_out)
begin
..
end
\end{lstlisting}
\end{lrbox}%
\scalebox{0.94}{\usebox{\mybox}}
\end{tabular}
\end{center}
Here, the traces $x$, $y$, $x'$ and $y'$ are given names $\mathtt{model\_in}$, $\mathtt{aux\_in}$, $\mathtt{model\_out}$, and $\mathtt{aux\_out}$, respectively.
The syntax for address $\mathtt{a}$ within trace $\mathtt{trace}$ is $\mathtt{trace[a]}$.
For example, to read the value of a continuous address $\mathtt{:}\mathtt{r}$ from the input model trace ($x$):
\noindent
\begin{center}
\begin{tabular}{c}
{\begin{lstlisting}[basicstyle=\small\ttfamily]
    val = @read(model_in[:r], :continuous)
\end{lstlisting}}
\end{tabular}
\end{center}
The syntax for writing to an address in $x' \oplus y' \in Z$ is similar.
For example to write a value $\mathtt{val}$ to $x'[\mathtt{:}\mathtt{r}]$:
\noindent
\begin{center}
\begin{tabular}{c}
{\begin{lstlisting}[basicstyle=\small\ttfamily]
    @write(model_in[:r], val, :continuous)
\end{lstlisting}}
\end{tabular}
\end{center}
Note that the input traces $x, y$ are distinct from the output traces $x', y'$;
input traces can only be read from, and output traces can only be written to.
For example, it is not possible to write an output trace and then read the written value from it later.
